On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-02-01 TTau_Halpha/submit/kurosawa.tex
Many typos fixed. (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@652 25464f21-6099-402c-b30a-640f1d40d38b

\documentclass[english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users
\newcommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

@@

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical
  T Tauri stars: the disc-wind-magnetosphere hybrid model}

@@

We present radiative transfer models of the circumstellar environment
of classical T~Tauri stars, concentrating on the formation of the
H$\alpha$ emission. The wide variety of line profiles seen in
observations are indicative of both inflow and outflow, and we
therefore employ a circumstellar structure that includes both
magnetospheric accretion and a disc wind.  We perform
systematic investigations of the model parameters for the wind and the
magnetosphere to search for possible geometrical and physical
conditions which lead to the types of profiles seen in
observations. We find that the hybrid models can reproduce the wide
range profile types seen is observations, and that the most common
profile types observed occupy a large volume of parameter
space. Conversely, the most infrequently observed profile morphologies
require a very specific set of models parameters.  We investigate the
wind contribution to the line profile as a function of model parameters, and
examine the reliability of H$\alpha$ as a mass accretion diagnostic.
Finally, we examine the
H$\alpha$ spectroscopic classification used by Reipurth et.~al, and
discuss the basic physical conditions that are required to reproduce
the profiles in each classified type.

@@

T~Tauri stars (TTS) are young ($\gtrsim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass objects, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some of
the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit excess continuum
flux in the ultraviolet (UV) and infrared (IR).  Their spectral energy
distribution and polarisation data suggest the presence of
circumstellar discs, which plays an important role in regulating
dynamics of gas flows around CTTS (e.g. \citealt{camenzind:1990}).

Many observational studies (e.g., \citealt{herbig:1962};
\citealt{edwards:1994}; \citealt{kenyon:1994};
\citealt*{reipurth:1996}; \citealt{alencar:2000}) of CTTS line
profiles have revealed evidence for both outward wind flows and inward
accretion flows, characterised by the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC)
profiles respectively. Typical mass-loss rates of CTTS are about
$10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kuhi:1964}; \citealt{edwards:1987};
\citealt*{hartigan:1995}), and the mass-accretion rates are also about
$10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988};
\citealt{gullbring:1998}).  Recent H$\alpha$ spectro-astrometric
observations by \citet*{takami:2003} provide indirect evidence for the
presence of bipolar and monopolar outflows down to $\sim
1\,\mathrm{au}$ scale (e.g.\,CS~Cha and RU~Lup).  Similarly, ESO VLT
observations using high-resolution ($R=50\,000$) two-dimensional
spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B, and HV~Tau~C) by
\citet{appenzeller:2005} show extended H$\alpha$ emission in the
direction perpendicular to the obscuring circumstellar disc,
suggesting the presence of the bipolar outflows. On an even larger
scale, \emph{HST} observations of HH30 \citep{burrows:1996} trace the
jet to within $\lesssim30$~au of the star. The jet has a cone shape
with an opening angle of $3^{\circ}$ between 70 and 700 au
\citep{koenigl:2000}. \citet{alencar:2000} found about 80 per cent of
their sample (30 CTTS) show blue-shifted absorption components in at
least one of the Balmer lines and Ca~K.

In the currently favoured model of accretion in CTTS, the accretion
disc is disrupted by the magnetosphere, which channels the gas from
the disc onto the stellar surface (e.g., \citealt{uchida:1985};
\citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported by recent measurements
of strong ($\sim 10^{3}\mathrm{G}$) magnetic fields in CTTS (e.g.,
\citealt{johns-krull:1999}; \citealt{symington:2005b}) and by
radiative transfer models which reproduce the gross characteristics of
observed profiles for some TTS (\citealt*{muzerolle:2001}). In
particular the magnetospheric accretion (MA) model explains blue-ward
asymmetric emission line profiles as resulting from the partial
occultation of the flow by the stellar photosphere, while inverse
P~Cygni profiles in the MA model result from inflowing material at
near free-fall velocities seen projected against hotspots on the
stellar surface.

Despite these successes, the overwhelming observational evidence for
outflow in the CTTS suggests that the MA model is only one component
of a complex circumstellar environment. Clearly, one must include the
contribution of any wind/jet flow if one wishes to both accurately
predict the mass-accretion rate and also determined the mass-loss rate
of CTTS via emission profile modelling. The first attempt in this
direction was made by \citet{alencar:2005} who demonstrated that the
observed H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better
reproduced by the radiative transfer model which included a collimated
disc-wind arising from near the inner edge of the accretion disc.

@@

In section~\ref{sec:model-configuration}, the model assumptions, and
the basic model configurations are presented. We discuss the radiative
transfer model used to compute the profiles in
section~\ref{sec:radiative-transfer-model}, and the results of model
calculations are given in section~\ref{sec:Results}.  We discuss our
results in the context of Reipurth's classification scheme in
section~\ref{sec:Discussion}, and our summary and conclusions are
presented in section~\ref{sec:Conclusions}.

@@

In order to understand how the different parts of the CTTS
circumstellar environment contribute to the formation of H$\alpha$,
the model space is divided into four different regions: (1)~a central
continuum source, (2)~the magnetospheric accretion flow, (3)~the wind
outflow, and (4)~the accretion disc. Fig.~\ref{fig:config_discwind}
depicts the relative location of the regions in the model space.  The
density is assumed to be rotationally symmetric around the
$z$-axis. The innermost radius of the magnetosphere at the equatorial
plane coincides with the the inner radius of the accretion disc. From
the innermost part of the accretion disc, the gas falls freely, moving
along the magnetic field onto the surface of the star.  In the
following subsections we describe the details of model components.

@@

  \caption{Basic model configuration of the disc-wind-magnetosphere
    hybrid model. The system consist of four components: (1)~the
    continuum source located at the origin of the
    cartesian coordinates $\left(x,y,z\right)$ -- the $y$-axis is into
    the paper, (2)~the magnetospheric accretion flow, (3)~the disc
    wind, and (4)~the accretion disc. The disc wind originates from
    the disc surface between $w_{i}=R_{\mathrm{wi}}$ and
    $w_{i}=R_{\mathrm{wo}}$ where $w_{i}$ is the distance from the $z$
    axis on the equatorial plane.  The wind source
    points ($S$), from which the stream lines diverges, are placed at
    distance $d$ above and below the star. The degree of wind
    collimation is controlled by changing the value of $d$.}

@@

We adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, \mathrm{R_{\sun}}$, $0.5\,
\mathrm{M_{\sun}}$, and $4000\,\mathrm{K}$ 
respectively. The model atmosphere of \citet{kurucz:1979} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) defines
the photospheric contribution to the continuum flux. The
parameters are summarised in Table~\ref{tab:std_parameters}.

An additional continuum source is considered for models which includes
magnetospheric accretion: as the infalling gas approaches the stellar
surface, it decelerates in a strong shock, and is heated to $\sim
10^{6}\,\mathrm{K}$.  The X-ray radiation produced in the shock will
be absorbed by the gas locally, and re-emitted in optical and UV
radiation (\citealt{koenigl:1991}; \citealt*{hartmann:1994}). This
will create hot rings where the magnetic field intersects with the
surface. We assume that the free-falling kinetic energy is thermalized
in the radiating layer, and is re-emitted as blackbody radiation with
a single temperature. With the parameters of the magnetosphere and the
star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate
is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of this accretion
luminosity to the photospheric luminosity is about $0.5$, and the
corresponding temperature of the hot rings is about
$6400\,\mathrm{K}$.  The continuum emission from the hot rings is
taken into account when computing the line profiles.

@@

The magnetic field and the gas stream lines are assumed to have the
following simple form:
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Fig.~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by
\citet{hartmann:1994} is adopted here. They computed the temperature,
assuming a volumetric heating rate which is proportional to $r^{-3}$,
by solving the energy balance of the radiative cooling rate of
\citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
\citet{martin:1996} presented a self-consistent determination of the
thermal structure of the inflowing gas along the dipole magnetic field
(equation~\ref{eq:dipole}) by solving the heat equation coupled to the
rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung
radiation and line emission from Ca~II and Mg~II ions. The results of
\citet{martin:1996} qualitatively agree with that of
\citet{hartmann:1994}.

@@

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}. Clearly there are several alternative outflow scenarios. Here we adopt
a simple kinematical model, based on the disc-wind paradigm, which broadly represents the results of
MHD simulations,

\citet*{knigge:1995} introduced the {}``split-monopole'' kinematic
disc-wind model in their studies of the UV resonance lines formed
in the winds of cataclysmic variable stars. Their formalism provides
a simple parameterisation of a disc wind that has similar properties 
to those found by  MHD modelling. In this model, the outflow
arises from the surface of the rotating accretion disc, and has a
biconical geometry. The specific angular momentum is assumed to be
conserved along a stream line, and the poloidal velocity component
is assumed to be simply a radial from vertically displaced {}``sources''
from the central star.  Here we briefly describe the disc-wind model, and
readers are referred to \citet{knigge:1995} and \citet{long:2002} for
details.

The four basic parameters of the model are: (1)~the mass-loss rate,
(2)~the degree of the wind collimation, (3)~the velocity gradient, and
(4)~the wind temperature. The basic configuration of the disc-wind
model is shown in Fig.~\ref{fig:config_discwind}. The disc wind
originates from the disc surface, but the {}``source'' point ($S$),
from which the stream lines diverges, are placed at distance $d$ above
and below the centre of the star.  The angle of the mass-loss
launching from the disc is controlled by changing the value of
$d$. The mass-loss launching occurs between $R_{\mathrm{wi}}$ and
$R_{\mathrm{wo}}$ where the former is set to be equal to the outer
radius of the magnetosphere ($R_{\mathrm{mo}}$) and the latter is set
to $1$~au as in \citet{krasnopolsky:2003}.

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       \left[ T\left(w\right) \right]  ^{4\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=4\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-5/2$, the
value of $\alpha$ is set to 0.76. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and the
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

The azimuthal/rotational component of the wind velocity
$v_{\phi}\left(w,z\right)$ is computed from the Keplerian rotational
velocity at the emerging point of the stream line
i.e. $v_{\phi}\left(w_{i},0\right)=\left(GM_{*}/w_{i}\right)^{1/2}$
where $w_{i}$ is the distance from the rotational axis ($z$) to the
emerging point on the disc, and by assuming the conservation of the
specific angular momentum along a stream line:
\begin{equation}
  v_{\phi}\left(w,z\right)\, = 
  v_{\phi}\left(w_{i},0\right)\,\left(\frac{w_{i}}{w}\right)\,\,.
  \label{eq:discwind-toroidal-velocity}
\end{equation}
Based on the canonical $\beta$ velocity law of hot stellar winds
(c.f.~\citealt*{castor:1975}), the poloidal component of the wind
velocity ($v_{p}$) parameterised as:
\begin{equation}
  v_{p}\left(w_{i},l\right) = c_{\mathrm{s}}\left(w_{i}\right) +
  \left[f\, v_{\mathrm{esc}}-c_{\mathrm{s}}\left(w_{i}\right)\right]\left(1-\frac{R_{s}}{l+R_{s}}\right)^{\beta}
  \label{eq:discwind-poloidal-velocity}
\end{equation}  
where $c_{\mathrm{s}}$, $f$, and $l$ are the sound speed at the wind
launching point on the disc, the constant scale factor of the
asymptotic terminal velocity to the local escape velocity (from the
wind emerging point on the disc), and the distance from the disc
surface along stream lines respectively. $R_{s}$ is the wind scale
length, and its value is set to 10~$R_{\mathrm{mi}}$ by following
\citet{long:2002}).

@@

  \caption{The dependency of the disc-wind density and velocity on the wind
    acceleration parameter $\beta$. The wind density $\rho$ (left panel)
    and the poloidal velocity component $v_{p}$ (right panel) along the
    stream line starting from the mid point of the wind launching zone,
    i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
    where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
    \right ) / 2$, are shown as a function of the distance ($l$) from the
    wind launching point (c.f.~
    equations~\ref{eq:discwind-poloidal-velocity} and
    \ref{eq:discwind-density}). The azimuthal velocity component
    ($v_{\phi}$), which is independent of $\beta$
    (c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
    the right panel for comparison. The density is normalised to the
    density $\rho_{0}$ at the wind launching point for the $\beta = 1.0$ case.
    The $v_{p}$ reaches the terminal velocity by 100~au for all
    $\beta$. In the far field ($l>10$~au), the density is approximately
    proportional to $\sim l^2$. }

>>>>

\documentclass[english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{babel}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users

%% Because html converters don't know tabularnewline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}
\providecommand{\tabularnewline}{\\}

\newcommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

@@

\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical
  T Tauri stars}

@@

We present radiative transfer models of the circumstellar environment
of classical T~Tauri stars, concentrating on the formation of the
H$\alpha$ emission. The wide variety of line profiles seen in
observations are indicative of both inflow and outflow, and we
therefore employ a circumstellar structure that includes both
magnetospheric accretion and a disc wind.  We perform
systematic investigations of the model parameters for the wind and the
magnetosphere to search for possible geometrical and physical
conditions which lead to the types of profiles seen in
observations. We find that the hybrid models can reproduce the wide
range of profile types seen in observations, and that the most common
profile types observed occupy a large volume of parameter
space. Conversely, the most infrequently observed profile morphologies
require a very specific set of models parameters.  We find our model
profiles are consistent with the canonical value of the 
mass-loss rate to mass-accretion rate ratio ($\mu=0.1$) found in
earlier magneto-hydrodynamic calculations and observations, but the
models with $0.05<\mu<0.2$ are still in accord with observed
H$\alpha$ profiles.  We investigate the
wind contribution to the line profile as a function of model parameters, and
examine the reliability of H$\alpha$ as a mass accretion diagnostic.
Finally, we examine the
H$\alpha$ spectroscopic classification used by Reipurth et.~al, and
discuss the basic physical conditions that are required to reproduce
the profiles in each classified type.

@@

T~Tauri stars (TTS) are young ($\gtrsim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass objects, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some of
the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit excess continuum
flux in the ultraviolet (UV) and infrared (IR).  Their spectral energy
distribution and polarisation data suggest the presence of
circumstellar discs, which play an important role in regulating
dynamics of gas flows around CTTS (e.g. \citealt{camenzind:1990}).

Many observational studies (e.g., \citealt{herbig:1962};
\citealt{edwards:1994}; \citealt{kenyon:1994};
\citealt*{reipurth:1996}; \citealt{alencar:2000}) of CTTS line
profiles have revealed evidence for both outward wind flows and inward
accretion flows, characterised by the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC)
profiles respectively. Typical mass-loss rates of CTTS are about
$10^{-9}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kuhi:1964}; \citealt{edwards:1987};
\citealt*{hartigan:1995}), and the mass-accretion rates are also about
$10^{-9}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988};
\citealt{gullbring:1998}).  Recent H$\alpha$ spectro-astrometric
observations by \citet*{takami:2003} provide indirect evidence for the
presence of bipolar and monopolar outflows down to $\sim
1\,\mathrm{au}$ scale (e.g.\,CS~Cha and RU~Lup).  Similarly, ESO VLT
observations using high-resolution ($R=50\,000$) two-dimensional
spectra of edge-on CTTS (HH~30, HK~Tau~B, and HV~Tau~C) by
\citet{appenzeller:2005} show extended H$\alpha$ emission in the
direction perpendicular to the obscuring circumstellar disc,
suggesting the presence of the bipolar outflows. On an even larger
scale, \emph{HST} observations of HH30 \citep{burrows:1996} trace the
jet to within $\lesssim30$~au of the star. The jet has a cone shape
with an opening angle of $3^{\circ}$ between 70 and 700 au
\citep{koenigl:2000}. \citet{alencar:2000} found about 80 per cent of
their sample (30 CTTS) show blue-shifted absorption components in at
least one of the Balmer lines and Na~D.

In the currently favoured model of accretion in CTTS, the accretion
disc is disrupted by the magnetosphere, which channels the gas from
the disc onto the stellar surface (e.g., \citealt{uchida:1985};
\citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported by recent measurements
of strong ($\sim 10^{3}\mathrm{G}$) magnetic fields in CTTS (e.g.,
\citealt{johns-krull:1999}; \citealt{symington:2005b}) and by
radiative transfer models which reproduce the gross characteristics of
observed profiles for some CTTS (\citealt*{muzerolle:2001}). In
particular the magnetospheric accretion (MA) model explains blue-ward
asymmetric emission line profiles as resulting from the partial
occultation of the flow by the stellar photosphere, while inverse
P~Cygni profiles in the MA model result from inflowing material at
near free-fall velocities seen projected against hotspots on the
stellar surface.

Despite these successes, the overwhelming observational evidence for
outflows in CTTS suggests that the MA model is only one component
of a complex circumstellar environment. Clearly, one must include the
contribution of any wind/jet flow if one wishes to both accurately
predict the mass-accretion rate and also determine the mass-loss rate
of a CTTS via emission profile modelling. The first attempt in this
direction was made by \citet{alencar:2005} who demonstrated that the
observed H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better
reproduced by the radiative transfer model which included a collimated
disc-wind arising from near the inner edge of the accretion disc.

@@

In Section~\ref{sec:model-configuration}, the model assumptions and
the basic model configurations are presented. We discuss the radiative
transfer model used to compute the profiles in
Section~\ref{sec:radiative-transfer-model}, and the results of model
calculations are given in Section~\ref{sec:Results}.  We discuss our
results in the context of Reipurth's classification scheme in
Section~\ref{sec:Discussion}, and our summary and conclusions are
presented in Section~\ref{sec:Conclusions}.

@@

In order to understand how the different parts of the CTTS
circumstellar environment contribute to the formation of H$\alpha$,
the model space is divided into four different regions: (1)~a central
continuum source, (2)~the magnetospheric accretion flow, (3)~the wind
outflow, and (4)~the accretion disc. Fig.~\ref{fig:config_discwind}
depicts the relative location of the regions in the model space.  The
density is assumed to be rotationally symmetric around the
$z$-axis. The innermost radius of the magnetosphere at the equatorial
plane coincides with the the inner radius of the accretion disc. From
the innermost part of the accretion disc, the gas falls freely, moving
along the magnetic field onto the surface of the star.  In the
following subsections we describe the details of the model components.

@@

  \caption{Basic model configuration of the disc-wind-magnetosphere
    hybrid model. The system consist of four components: (1)~the
    continuum source located at the origin of the
    cartesian coordinates $\left(x,y,z\right)$ -- the $y$-axis is into
    the paper, (2)~the magnetospheric accretion flow, (3)~the disc
    wind, and (4)~the accretion disc. The disc wind originates from
    the disc surface between $w_{i}=R_{\mathrm{wi}}$ and
    $w_{i}=R_{\mathrm{wo}}$ where $w_{i}$ is the distance from the $z$
    axis on the equatorial plane.  The wind source
    points ($S$), from which the stream lines diverge, are placed at
    distance $d$ above and below the star. The degree of wind
    collimation is controlled by changing the value of $d$.}

@@

We adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature of the photosphere
($T_{\mathrm{ph}}$) are $2\, \mathrm{R_{\sun}}$, $0.5\,
\mathrm{M_{\sun}}$, and $4000\,\mathrm{K}$ 
respectively. The model atmosphere of \citet{kurucz:1979} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) defines
the photospheric contribution to the continuum flux. The
parameters are summarised in Table~\ref{tab:std_parameters}.

An additional continuum source is considered for models which include 
magnetospheric accretion: as the infalling gas approaches the stellar
surface, it decelerates in a strong shock, and is heated to $\sim
10^{6}\,\mathrm{K}$.  The X-ray radiation produced in the shock will
be absorbed by the gas locally, and re-emitted  optical and UV
radiation (\citealt{koenigl:1991}; \citealt*{hartmann:1994}). This
will create hot rings where the magnetic field intersects with the
surface. We assume that the free-falling kinetic energy is thermalized
in the radiating layer, and is re-emitted as blackbody radiation with
a single temperature. With the parameters of the magnetosphere and the
star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate
is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of this accretion
luminosity to the photospheric luminosity is about $0.5$, and the
corresponding temperature of the hot rings is about
$6400\,\mathrm{K}$.  The continuum emission from the hot rings is
taken into account when computing the line profiles.

@@

The magnetic field and the gas stream lines are assumed to have the
following simple form 
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are the radial and
the polar component of a field point position vector in the accretion
stream.  $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$), and its value is
restricted between $R_{\mathrm{mi}}$ and 
$R_{\mathrm{mo}}$ (Fig.~\ref{fig:config_discwind}). 
Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by
\citet{hartmann:1994} is adopted here. They computed the temperature,
assuming a volumetric heating rate which is proportional to $r^{-3}$,
by solving the energy balance of the radiative cooling rate of
\citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
\citet{martin:1996} presented a self-consistent determination of the
thermal structure of the inflowing gas along the dipole magnetic field
(equation~\ref{eq:dipole}) by solving the heat equation coupled to the
rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung
radiation and line emission from Ca~II and Mg~II ions. The results of
\citet{martin:1996} qualitatively agree with those of
\citet{hartmann:1994}.

@@

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of the stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced (see also
\citealt{krasnopolsky:2003}).  Recent reviews on the jet/wind-disc
connection can be found in \citet{koenigl:2000} and
\citet{pudritz:2005}. Clearly there are several alternative outflow
scenarios. Here we adopt a simple kinematical model, based on the
disc-wind paradigm, which broadly represents the results of MHD
simulations,

\citet*{knigge:1995} introduced the {}``split-monopole'' kinematic
disc-wind model in their studies of the UV resonance lines formed
in the winds of cataclysmic variable stars. Their formalism provides
a simple parameterisation of a disc wind that has similar properties 
to those found by  MHD modelling. In this model, the outflow
arises from the surface of the rotating accretion disc, and has a
biconical geometry. The specific angular momentum is assumed to be
conserved along a stream line, and the poloidal velocity component
is assumed to be simply a radial from {}``sources'' vertically displaced 
from the central star.  Here we briefly describe the disc-wind model, and
readers are referred to \citet{knigge:1995} and \citet{long:2002} for
details.

The four basic parameters of the model are: (1)~the mass-loss rate,
(2)~the degree of the wind collimation, (3)~the velocity gradient, and
(4)~the wind temperature. The basic configuration of the disc-wind
model is shown in Fig.~\ref{fig:config_discwind}. The disc wind
originates from the disc surface, but the {}``source'' point ($S$),
from which the stream lines diverge, are placed at distance $d$ above
and below the centre of the star.  The angle of the mass-loss
launching from the disc is controlled by changing the value of
$d$. The mass-loss launching occurs between $R_{\mathrm{wi}}$ and
$R_{\mathrm{wo}}$ where the former is set to be equal to the outer
radius of the magnetosphere ($R_{\mathrm{mo}}$) and the latter is set
to $1$~au,  as in \citet{krasnopolsky:2003}.

The local mass-loss rate per unit area ($\dot{m}$) is assumed to
be proportional to the mid-plane temperature of the disc, and is a
function of the cylindrical radius $w=\left(x^{2}+y^{2}\right)^{1/2}$, i.e.
\begin{equation}
  \dot{m}\left(w\right) \propto
       \left[ T\left(w\right) \right]  ^{4\alpha}\,\,.\label{eq:discwind-massloss-temp}
\end{equation}
The mid-plane temperature of the disc is assumed to be expressed
as a power-law in $w$; thus, $T\propto w^{q}$. Using this in the
relation above, one finds 
\begin{equation}
  \dot{m}\left(w\right)\propto w^{p}
  \label{eq:discwind-massloss-w}
\end{equation}
where $p=4\alpha\times q$. The index of the mid-plane temperature power
law is adopted from the dust radiative transfer model of
\citet{whitney:2003a} who found the innermost part of the accretion
disc has $q=-1.15$.  In order to be consistent with the collimated
disc-wind model of \citet{krasnopolsky:2003} who used $p=-5/2$, the
value of $\alpha$ is set to 0.76. The constant of proportionality in
equation~\ref{eq:discwind-massloss-w} is found by integrating
$\dot{m}$ from $R_{\mathrm{wi}}$ to $R_{\mathrm{wo}}$, and 
normalising the value to the total mass-loss rate
$\dot{M}_{\mathrm{wind}}$.

The azimuthal/rotational component of the wind velocity
$v_{\phi}\left(w,z\right)$ is computed from the Keplerian rotational
velocity at the emerging point of the stream line
i.e. $v_{\phi}\left(w_{i},0\right)=\left(GM_{*}/w_{i}\right)^{1/2}$
where $w_{i}$ is the distance from the rotational axis ($z$) to the
emerging point on the disc, and by assuming the conservation of the
specific angular momentum along a stream line:
\begin{equation}
  v_{\phi}\left(w,z\right)\, = 
  v_{\phi}\left(w_{i},0\right)\,\left(\frac{w_{i}}{w}\right)\,\,.
  \label{eq:discwind-toroidal-velocity}
\end{equation}
Based on the canonical $\beta$ velocity law of hot stellar winds
(c.f.~\citealt*{castor:1975}), the poloidal component of the wind
velocity ($v_{p}$) parameterised as:
\begin{equation}
  v_{p}\left(w_{i},l\right) = c_{\mathrm{s}}\left(w_{i}\right) +
  \left[f\, v_{\mathrm{esc}}-c_{\mathrm{s}}\left(w_{i}\right)\right]\left(1-\frac{R_{s}}{l+R_{s}}\right)^{\beta}
  \label{eq:discwind-poloidal-velocity}
\end{equation}  
where $c_{\mathrm{s}}$, $f$, and $l$ are the sound speed at the wind
launching point on the disc, the constant scale factor of the
asymptotic terminal velocity to the local escape velocity (from the
wind emerging point on the disc), and the distance from the disc
surface along stream lines respectively. $R_{s}$ is the wind scale
length, and its value is set to 10~$R_{\mathrm{mi}}$ by following
\citet{long:2002}.

@@

  \caption{The dependency of the disc-wind density and velocity on the wind
    acceleration parameter $\beta$. The wind density $\rho$ (right panel)
    and the poloidal velocity component $v_{p}$ (left panel) along the
    stream line starting from the mid point of the wind launching zone,
    i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
    where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
    \right ) / 2$, are shown as a function of the distance ($l$) from the
    wind launching point (c.f.~
    equations~\ref{eq:discwind-poloidal-velocity} and
    \ref{eq:discwind-density}). The azimuthal velocity component
    ($v_{\phi}$), which is independent of $\beta$
    (c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
    the right panel for comparison. The density is normalised to the
    density $\rho_{0}$ at the wind launching point for the $\beta = 1.0$ case.
    The $v_{p}$ reaches the terminal velocity by 100~au for all
    $\beta$. In the far field ($l>10$~au), the density is approximately
    proportional to $\sim l^2$. }
